---
title: "Data Wrangling and Analysis in R"
output:
  html_document:
    df_print: paged
---
  
## 1. Installing packages
  
```{r}
install.packages("tidyverse")
```

#### Load tidyverse for use in our notebook
```{r}
library(tidyverse)
```
## 2. Data Importation
```{r}
mtn_df <- read_csv("https://bit.ly/2ZlpzjF")
mtn_df
```
## 3. Data Understanding
```{r}
#preview the first 6 records of the mtn dataset
head(mtn_df)
```
```{r}
#preview the last 6 records of the mtn dataset
tail(mtn_df)
```

```{r}
library(dplyr)
```
```{r}
#display the variable names, datatypes and first few records of the mtn dataset
glimpse(mtn_df)
```
```{r}
#display more information about the datatypes
str(mtn_df)
```
```{r}
#check the size of the dataset
dim(mtn_df)
```
## 4. Finding and dealing with Duplicates 
```{r}
mtn_df[duplicated(mtn_df),]
```
```{r}
#We remove duplicates based on customerId column
mtn_df[!duplicated(mtn_df$customerID), ]

```
## 5. Missing data
```{r}
#Lets try to identify missing data and pass the dataframe
is.na(mtn_df)
```
```{r}
#we find the total missing values in each column
colSums(is.na(mtn_df))
```
###We have severalmissing data in all columns except the customerID
### We drop all rows containing missing values.
```{r}
na.omit(mtn_df)
```
```{r}
#we replace the missing values in monthly charge and totalcharge with the mean
mtn_df$MonthlyCharges[is.na(mtn_df$MonthlyCharges)] <- mean(mtn_df$MonthlyCharges, na.rm = TRUE)
mtn_df$TotalCharges[is.na(mtn_df$TotalCharges)] <- mean(mtn_df$TotalCharges, na.rm = TRUE)
mtn_df
```
## 6. Find and Dealing with Outliers
```{r}
head(mtn_df)
```

```{r}
# Calculating lower and upper percentiles per MonthlyCharges
lower_bound <- quantile(mtn_df$MonthlyCharges, 0.025)
upper_bound <- quantile(mtn_df$MonthlyCharges, 0.975)
lower_bound
upper_bound
```

###Lower bound for Monthly charges is 19.4 and upper bound is 111.05 
```{r}
# Calculating lower and upper percentiles per TotalCharges
lower_bound <- quantile(mtn_df$TotalCharges, 0.025)
upper_bound <- quantile(mtn_df$TotalCharges, 0.975)
lower_bound
upper_bound
```
###Lower bound for Total charges is 24.9225 and upper bound is 7558.954  

## 7. Filtering and sorting
```{r}
head(mtn_df)
```
```{r}
#view customers with the highest tenure and gender
mtn_df %>% 
  select(GENDER,SeniorCitizen, tenure) %>%
  arrange(desc(tenure))
```
###From the analysis above the people with the highesttenure above 100 were majorly male and were not senior citizens. The female who feature in that caegory are both senior citizens
```{r}
mtn_df %>%
  group_by(GENDER) %>%
  summarize(TotalCharges = mean(TotalCharges))
```
###Male Gender contributed to highest charges from the analysis above
```{r}
selected_df <- filter(mtn_df, Churn == "Yes")
select(selected_df, PARTNER, Churn)
```
### We notice that both partners and mtn customers were churning
```{r}
selected_df %>%
  arrange(desc(tenure))
```
```{r}
selected_df %>%
  group_by(GENDER) %>%
  summarize(count = sum(customerID))
```